---
title: "Bellabeat Capstone"
author: "Sara Romero"
date: "2023-12-06"
output:
  pdf_document: default
  html_document: default
Email: SaraRomero43@gmail.com
LinkedIn: linkedin.com/in/sararomero43/
---
#Introduction
As part of my Google Data Analytics Certificate capstone project, I assumed the role of a marketing analyst, focusing on Bellabeat—a technology manufacturer specializing in women's health products.

#Questions
what are some trends in personal health tracker usage?
How can we apply these trends to Bellabeat?

#prepare
30 participants generated 18 data sets, which were gathered through Amazon Mechanical Turk from December 3, 2016, to December 5, 2016.

LIMITATION
This data lacks demographic details, potentially limiting its applicability to Bellabeats' female customer base. Furthermore, the small sample size of only 30 participants poses challenges for conducting a comprehensive analysis.

#process
I chose to use 3 of the 18 data sets:
  dailyActivity_merged
  SleepDay_merged
  weightLoginfo_merged
These particular datasets were picked due to their comprehensive inclusion of information found in the other sets.

I used Google sheets to clean by:
  check for duplicates 
  trim white space
  formatted the date and time into just the date

#Analyze & Share

##Importing data sets
 Only using 3 of the 18 sets that are available as they have everything needed.

```{r}
DailyActivity <- read_excel("D:/data analyst/projects/BellaBeat/BellabeatData11.23.xlsx", 
sheet = "dailyActivity_merged.csv")

Sleep <- read_excel("D:/data analyst/projects/BellaBeat/BellabeatData11.23.xlsx", 
sheet = "sleepDay_merged")

WeightLog <- read_excel("D:/data analyst/projects/BellaBeat/BellabeatData11.23.xlsx", 
sheet = "weightLogInfo_merged")
```
##using 
  comparing fairly/very active against light/sedentary activity by changing the minutes to hours and then putting them in a table.

started by getting the minutes switched to hours and placing it in a data frame. 
```{r}
SedentaryHour <- DailyActivity$SedentaryMinutes/60
LightlyActiveHour <- DailyActivity$LightlyActiveMinutes/60
FairlyActiveHour <- DailyActivity$FairlyActiveMinutes/60
VeryActiveHour <- DailyActivity$VeryActiveMinutes/60

HourlyActivity <- data.frame(SedentaryHour, LightlyActiveHour, FairlyActiveHour, VeryActiveHour)

View(HourlyActivity)
```
###merging HourlyActivity into DailyActivity
```{r}
AllActivity <- merge(HourlyActivity, DailyActivity, Id="Id" , all.HourlyActivity=TRUE)

View(AllActivity)
``` 
##

```{r}
DA_grp_ID = AllActivity %>%
  group_by(Id) %>%
  summarise(Light_sedentary = sum(SedentaryHour, 
            LightlyActiveHour),
            Very_Fairly = sum(FairlyActiveHour,                      VeryActiveHour),
            .groups = 'drop')


View(DA_grp_ID)
```


Just splitting it up so that it makes the chart THEN you can view it.

```{r}
view(DA_grp_ID)
```

## creating a graph for DA_grp_ID
  This is in minuets 
```{r}
ggplot(data = DA_grp_ID, mapping = aes(fill = Light_sedentary, factor(Very_Fairly), y= Light_sedentary)) +
  geom_bar(width = .85, position = "stack", stat= "identity") +
  facet_grid(scales = "free") +
  theme(axis.text.x = element_text(angle=90, vjust=5, hjust=1))
```
##Creating bar graph  
  I want to see if I can find any trends between Total steps and Calories.

```{r}
ggplot(data = DailyActivity, mapping = aes(x=Calories, y=TotalSteps)) +
  geom_bar(color= 'blue', stat = 'identity')
  
```

##Creating stacked bar graph
  Looking at the total time in bed verses total time asleep. Data is in minutes so 480mins = 8hrs.

```{r}
ggplot(data = Sleep, mapping = aes(x=TotalTimeInBed, y=TotalMinutesAsleep)) +
  geom_point()
```
